---
title: "Calendar Project"
author: "Majd Rouhana"
date: "9/10/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(lubridate)
library(ical)
library(ggplot2)
library(ggformula)
library(stringr)
library(anytime)
```
```{r}
path <- "/home/class22/mrouhana22/git/majd_STAT231/projects"
filename <- "STAT231_c_pl4hgkq7kd4cehd0u83k85aja4@group.calendar.google.com.ics"

my_calendar0 <- ical_parse_df(file = paste0(path,"/",filename)) %>%
  mutate(start_datetime = with_tz(start, tzone = "America/New_York")
    , end_datetime = with_tz(end, tzone = "America/New_York")
    , length_hour = end_datetime - start_datetime
    , date = floor_date(start_datetime, unit = "day")
    , day = weekdays(as.Date(date))
    , start_time = strftime(start, format="%H:%M:%S"))
```
```{r}
my_calendar0 %>%
  group_by(summary, length_hour) %>%
  summarize(count = n())

my_calendar0 %>%
  summarize(tot_hrs = as.numeric(sum(length_hour)))

sleepCalendar <- my_calendar0 %>%
  filter(summary == "Sleep") %>%
  group_by(day) %>%
  summarize(meanSlp = round(mean(length_hour), digits = 2))

ggplot(data = sleepCalendar, aes(x = day, y = meanSlp, fill = day)) + geom_bar(stat = "identity")

eatCalendar <- my_calendar0 %>%
  mutate(eat = ifelse((summary == "Coffee" | summary == "Lunch" | summary == "Dinner" | summary == "Snack"),
                      "Eating", "Not Eating"),
         numDay = ifelse((day == "Friday" | day == "Saturday"), 3, 2)) %>%
  filter(eat == "Eating") %>%
  subset(select = c(eat, day, numDay)) %>%
  group_by(day) %>%
  summarize(meanEat = round(n() / numDay, digits = 2)) %>%
  arrange(meanEat) %>%
  slice(from = 1, to = 12)

ggplot(data = eatCalendar, aes(x = day, y = meanEat, fill = day)) + geom_bar(stat = "identity")

musicCalendar <- my_calendar0 %>%
  mutate(numDay = ifelse((day == "Friday" | day == "Saturday"), 3, 2)) %>%
  filter(summary == "Music") %>%
  group_by(day) %>%
  summarize(meanPractice = round(n() / numDay))

tableCalendar <- merge(sleepCalendar, eatCalendar)

knitr::kable(tableCalendar, col.names = c("Day of the Week",
             "Average Hours of Sleep", "Average Number of Times I Ate"))



sleepCalendar2 <- my_calendar0 %>%
  filter(summary == "Sleep") %>%
  mutate(sleepHr = length_hour)

eatCalendar2 <- my_calendar0 %>%
  mutate(eat = ifelse((summary == "Coffee" | summary == "Lunch" |
                       summary == "Dinner" | summary == "Snack"),
                       "Eating", "Not Eating"),
         numDay = ifelse((Day == "Friday" | Day == "Saturday"), 3, 2)) %>%
  filter(eat == "Eating") %>%
  group_by(Day) %>%
  summarize(count = n())

mergedCalendar2 <- merge(sleepCalendar2, eatCalendar2)
```
In the thirteen days logged so far, I slept and ate lunch and dinner every day (thankfully). On average, I slept about 7.6 hours a night. On five of those days I drank coffee, while on six of those days I ate a snack. I also practiced music on eight of the thirteen days.

I am looking to see if I can create a table that has sleep on the left side, with the other summaries on the top. Maybe I can use the `gather()` and `spread()` functions to achieve this. I want to see the number of times I drink coffee, eat lunch, dinner or a snack, or play my instrument depending on the number of hours I sleep. For example, do I practice more often on days I sleep more? Do I consume calories more on days I sleep less?